---
title: "Lab1_Morrison"
author: "Grant Morrison"
date: "4/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r, message = FALSE}
library(sf)
library(tmap)
library(tidyverse)
library(ggplot2)
```


Reading data from the internet using **sf** function `st_read`
```{r}
chicago <- st_read("https://data.cityofchicago.org/resource/igwz-8jzy.geojson")
```


Plotting geometry column of the chicago data set with `plot`
```{r}
plot(chicago$geometry)
```


Setting tmap mode to `"view"` with `tmap_mode` and plotting the community areas
with `tm_shape` and `tm_borders`
```{r}
tmap_mode("view")
tm_shape(chicago) +
  tm_borders()
```


Using **tidyverse** function `filter` to get the Hyde Park community area
```{r}
hyde_park <- chicago %>% filter(community == "HYDE PARK")
```



Plotting the community areas with Hyde Park as blue with **tmap** functions
```{r}
tm_shape(chicago) +
  tm_borders() +
  tm_shape(hyde_park) +
  tm_fill(col = "blue")
```


Computing centroids of Chicago community areas with `st_centroid`
```{r, warning=FALSE}
centroids <- st_centroid(chicago)
```


Ploting centroids in red, Chicago community areas, and Hyde Park
in Blue with **tmap** functions
```{r}
tm_shape(chicago) +
  tm_borders() +
  tm_shape(hyde_park) +
  tm_fill(col = "blue") +
  tm_shape(centroids) +
  tm_dots(col = "red")
```









