---
title: "Lab2_Morrison"
author: "Grant Morrison"
date: "4/18/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message = FALSE}
library(tmap)
library(sf)
library(ggplot2)
library(tidyverse)
library(geodaData)
```



## Geoda Data

Data is loaded from the geodaData package, which contains the datasets used in the
Geoda workbook. The data is demographic and housing data for NYC's suboroughs.
```{r}
nyc <- geodaData::nyc_sf
```

## Summary Statistics


**rentpct08** is the percentage of housing stock that is market rate rental units.
`summary` gives basic summary statistics for the variable **rentpct08**
```{r}
summary(nyc$rentpct08)
```

`sd` gives the standard deviation of **rentpct08**
```{r}
sd(nyc$rentpct08)
```


We do the same thing for **hhsiz08**, which is the average number of people per 
household.
```{r}
summary(nyc$hhsiz08)
```


```{r}
sd(nyc$hhsiz08)
```

## Histograms

Using **ggplot2** we can get a histogram of **rentpct08**
```{r}
ggplot(data = nyc, aes(x=rentpct08)) +
  geom_histogram(bins = 30)
```


Now we make a histogram for **hhsiz08**
```{r}
ggplot(data = nyc, aes(x=hhsiz08)) +
  geom_histogram(bins = 30)
```


## Mapping

We map **rentpct08** with **tmap** functions. `tmap_mode` gives us a 
basemap. 
```{r}
tmap_mode("view")
tm_shape(nyc) +
  tm_fill("rentpct08", style = "quantile") +
  tm_borders() 
```


Here we map **hhsiz08**, but with a blue palette.
```{r}
tm_shape(nyc) +
  tm_fill("hhsiz08", style = "quantile", palette = "Blues") +
  tm_borders() 
```
