---
title: "Lab3_Morrison"
author: "Grant Morrison"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, message=FALSE}
library(sf)
library(tidyverse)
library(tmap)
library(RSocrata)
library(ggplot2)
```


Reading data from socrata portal on the Chicago Open Data Portal website. Data is crimes
for the past year to date. `read.socrata` reads data from socrata data APIs on the web.
```{r}
crimes <- read.socrata("https://data.cityofchicago.org/resource/x2n5-8w5q.csv")
```



Reading the community area shapefile from the web with `st_read`
```{r}
chicago <- st_read("https://data.cityofchicago.org/resource/igwz-8jzy.geojson")
```

Plotting the Chicago community areas
```{r}
plot(chicago$geometry)
```


Checking the different types of crime with `unique`
```{r}
unique(crimes$X_primary_decsription)
```


Filtering out all crimes that are not homicides with `filter`.
```{r}
homicides <- crimes %>% filter(X_primary_decsription == "HOMICIDE")
```

Checking the number of homicides with `nrow`
```{r}
nrow(homicides)
```


Converting the data frame to **sf** with `st_as_sf` and setting the crs with `st_crs`
```{r}
homicide_points <- st_as_sf(homicides, coords = c("longitude","latitude"))
st_crs(homicide_points) <- 4326
```





Changing the **tmap** mode to "view" for base map and interactivity
```{r}
tmap_mode("view")
```



Plotting the community areas and homicide points with **tmap** functions
```{r}
tm_shape(chicago) +
  tm_borders() +
  tm_shape(homicide_points) +
  tm_dots(col = "red")
```
I will aggregate the points to communtiy areas in order to make choropleth map



To start, I spatially join the community area identifier to the points.
```{r, message = FALSE}
comm.points <- st_join(homicide_points,chicago["area_num_1"])
```


**area_num_1** is a factor, so I change it to type integer with `as.integer` to avoid problems later.
```{r}
comm.points$area_num_1 <- as.integer(comm.points$area_num_1)
chicago$area_num_1 <- as.integer(chicago$area_num_1)
```

Next, I set the geometry of the points to NULL for convenience in making the counts.
```{r}
st_geometry(comm.points) <- NULL
```


`count` creates counts of the number of points with the same integer for **area_num_1**.`head` gives the
first few observations 
```{r}
homicide.cnts <- comm.points %>% count(area_num_1)
head(homicide.cnts)
```


I change the names to something more meaningful with `rename`
```{r}
homicide.cnts <- homicide.cnts %>% rename(comm = area_num_1, homicide_count = n)
head(homicide.cnts)
```


Now I join the counts to the original Chicago community area boundaries with `left_join`
```{r}
chicago <- left_join(chicago,homicide.cnts, by = c("area_num_1" = "comm"))
```



I plot the homicide counts for each community area with **tmap** functions
```{r}
tm_shape(chicago) +
  tm_fill("homicide_count", style = "quantile", palette = c("blue","white","red")) +
  tm_borders()
```


Here I assign 0 for missing values, so I can make a density plot. This is easily done with `is.na` and
bracket notation.
```{r}
chicago$homicide_count[is.na(chicago$homicide_count)] <- 0
```


Lastly with **ggplot2**, I make a density plot to view the univariate distribution of homicide counts
```{r}
ggplot(chicago, aes(x=homicide_count)) +
  geom_density()
```












