---
title: "Lab 6 GIS3"
author: "Grant Morrison"
output:
  pdf_document: default
  html_document: default
---

The data and some of the code were from the Urban Analytics workshop by Alex Singleton.
The github link to this workshop can be found at [Urban Analytics](https://github.com/alexsingleton/urban_analytics)

### Loading the libraries


```{r}
library(tmap)
library(sf)
```

### Loading the data

This data was created earlier in the lab, but we just load it here with `load` to keep things
simpler and cleaner. I went through the whole lab, but I didn't want to have all of Dr. Singleton's code 
copied into my lab on my repository. I shortened it to just loading the clusters and 
mapping them.
```{r}
load("./data/cluster_7.Rdata")
```

The cluster_7 object contains a list of different outputs related to the cluster analysis- we can view these:

```{r}
str(cluster_7)
```

The cluster results can therefore be accessed as follows:

```{r}
lookup <- data.frame(cluster_7$cluster)
lookup$OA <- rownames(lookup)
colnames(lookup) <- c("K_7","OA")
lookup$SUPER <- LETTERS[lookup$K_7]
```

We will also look at the distribution of these clusters:

```{r}
table(lookup$K_7)
```

To get the boundaries, we use `read_sf`, which also works on .geojson files.
```{r}
liverpool_sf = read_sf( "./data/Liverpool_OA_2011.geojson")
```

Here we merge the cluster data together with the boundary, by the OA code.

```{r}
liverpool_sf <- merge(liverpool_sf, lookup, by.x="oa_code",by.y="OA")
```

### Mapping the Clusters


To map the clusters, we use **tmap**. First we set `tmap_mode("view")`. Then we use
`tm_fill` to map the cluster with palette **Set2**. The we add black borders with
`tm_borders`. With `tm_layout`, we set the title to be "Cluster Map Liverpool"
```{r}
tmap_mode("view")
tm_shape(liverpool_sf) + 
  tm_fill("SUPER", palette = "Set2") +
  tm_borders(col = "black",alpha =.5) +
  tm_layout(title = "Cluster Map Liverpool")
```

The interpretation for this map is fairly difficult as the clustering is done with many variables. The
variables used are categorized as demographic, socioeconomic, and housing. SMR illness rates are computed to add to 
the cluster analysis. Each variable is standardized to a range between 0 and 1, so that each variable
affects the cluster analysis. The cluster groups are spherical clusters that closest similarity
with respect to the input variables. I am not familiar with Liverpool as a city, but it looks like group B is most
common throughout the map.

